Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities

### 2021-03-16 writing/transect_paper_rev2.tex
output diff from second revision (to make sure there are no regressions from accepted version)

Although ECM fungi form many well-known mushrooms (e.g., \emph{Amanita}, \emph{Cantharellus}, \emph{Boletus}), some instead produce inconspicuous (e.g., \emph{Tomentella}) or no (e.g., \emph{Cenococcum}) fruit bodies.
Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by sampling of vegetative structures \autocite{horton2001}.
Unlike many saprotrophic fungi which grow easily in axenic culture, ECM fungi are usually difficult to culture, so DNA barcoding is increasingly used to investigate vegetative structures in the field.
The advent of high-throughput sequencing (HTS) has facilitated such studies by providing enough sequencing depth for metabarcoding of bulk environmental samples such as soils \autocite{Lindahl2013}.

As additional techniques and methods are developed for HTS, there is an increasing array of choices for researchers investigating fungal communities.
Fungal metabarcoding studies using short-read HTS technologies such as 454 Pyrosequencing, Illumina, and Ion Torrent
have usually targeted the rDNA internal transcribed spacer regions ITS1 or ITS2,
which are the standard molecular barcode for fungi, providing sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS \autocite{schoch2012,Lindahl2013}.
In some groups such as arbuscular mycorrhizal fungi, variable regions of the rDNA small subunit (SSU) are the barcode of choice \autocite{opik2010}, and variable regions of the rDNA large subunit (LSU) have also been used for barcoding \autocite{house2016,tedersoo2015}.
The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis \autocite{Lindahl2013}.
If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST \autocite{altschul1990,Lindahl2013} with public databases such as GenBank \autocite{benson2013} and Unite \autocite{nilsson2019a}.
However, this approach comes with some potential weaknesses, discussed below.

Different sequencing technologies have different capabilities in terms of sequencing depth and read length, as well as differing quality profiles and potential biases \autocite{yang2013}.
The rapid development of new HTS technologies, as well as subsequent iterative improvements in sequencing chemistry and read capacity, means that the technologies used in metabarcoding studies, along with any associated biases, change frequently.
As an example, the first study using HTS metabarcoding of soil fungi was published in 2009 \autocite{buee2009} using 454 Pyrosequencing; production of 454 sequencers was subsequently discontinued in 2015, and sales of reagents stopped in 2016 \autocite{hollmer2013}.
This brings into question the comparability of studies conducted only a few years apart.

ITS1 and ITS2 often have suitable variation to distinguish species, although closely related species may share identical ITS sequences in certain groups such as various Pezizomycotina \autocite{schoch2012},
but this variablity means that they cannot be reliably aligned over the fungal kingdom \autocite{Lindahl2013,Tedersoo2018}.
Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa \autocite{ihrmark2012,tedersoo2015,palmer2018}.
Further bias is introduced by variation in the 5.8S region which separates the two ITS regions, as well as in the 5' end of LSU, which makes it difficult to design primers that are suitable for all fungi \autocite{tedersoo2015}.

Distance-based clustering conflates intra-species variation and sequencing error, and results are dataset-specific.
In contrast, more recent denoising methods such as DADA2 \autocite{callahan2017}, Deblur \autocite{amir2017}, and UNOISE2 \autocite{edgar2016} utilize read quality information to control for sequencing error while preserving intra-species variation.
The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent true amplicon sequences from the sample.
Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, although alpha and beta diversity estimates based on ASVs and OTUs at different clustering thresholds are highly correlated \autocite{glassman2018,botnen2018}.
ASVs have been suggested to be less dataset specific that cluster-based OTUs \autocite{callahan2017}.
Support for PacBio has recently been added to DADA2 \autocite{callahan2019}, but its application requires greater sequencing depth for longer reads, especially in high diversity samples.

Because both OTU clustering and denoised ASVs may ``clump'' different species into a single unit and ``split'' a single species into multiple units \autocite{ryberg2015}, diversity measures based on counting species within a community or shared species between two communities may give different results depending on the clustering threshold.
In contrast, phylogenetic community distance measures \autocite{wong2016} are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree \autocite{matsen2010,berger2011}, but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments \autocite{tedersoo2018}, or to create hybrid trees using ITS and a more conserved marker such as SSU or LSU based on matching taxonomic annotations in reference databases \autocite{fouquier2016}, but these approaches are only applicable to sequences of known taxonomic affiliation.

Assignment of taxonomic identities to environmental sequences is dependent on both the reference database and the algorithm used.
Although the public INSDC databases \autocite{karsch-mizrachi2018} are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations \autocite{nilsson2006,steinegger2020} as well as sequences of poor technical quality \autocite{nilsson2012}.
Consequently taxonomic assignments based on these databases may be incorrect or inconsistent.
Several curated databases also exist which attempt to address these issues and which cover the whole fungal kingdom.
The Unite database is a more curated attempt to include all publically available ITS sequences (800~000 as of release 8.0), originally targetting fungi but now expanded to include all eukaryotes \autocite{nilsson2019a}, where efforts have been made to correct incorrect annotations and exclude low-quality sequences \autocite{abarenkov2018a}.
The Ribosomal Data Project \autocite[RDP,][]{cole2014} hosts two manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
the Warcup ITS database, containing 18~000 manually curated fungal ITS sequences \autocite{deshpande2016}, and the RDP fungal LSU training set, containing 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups \autocite{liu2012}.
Although the quality of sequences and taxonomic annotations is undoubtedly higher in these more curated databases, they are inherently limited in taxonomic coverage and do not include the most recently published sequences.

@@

Recent long-read HTS technologies such as Pacific Biosciences (PacBio) Single Molecule Real Time (SMRT) sequencing enable sequencing longer amplicons which include both the ITS regions and the flanking, more highly conserved SSU and/or LSU regions.
This can improve taxonomic placement of sequences that lack close database matches and allow the alignment of metabarcoding reads for subsequent phylogenetic analysis \autocite{Tedersoo2018}.
Information from phylogenetic trees produced from long-amplicon metabarcoding has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number \autocite{Kennedy2018}.

Because of the variety of sequencing platforms and analytical pipelines which have been used in metabarcoding studies, comparisons between studies may be difficult.
Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
Knowledge of turnover scale is important when planning studies of local diversity and its environmental correlates.
Turnover scale varies between different ecosystems and taxonomic groups,
and can be measured by the range at which a Mantel correlogram indicates significant autocorrelation,
or by fitting a function to an empirical distance-decay curve of community dissimilarity vs.~distance \autocite{legendre2012}.

We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDP classifier, SINTAX, IDTAXA)
with three different reference databases (Unite, Warcup, RDP),
and both non-phylogenetic and phylogenetic community distance measures.
We also present new algorithms for dividing the LSU into domains, combining denoising results from multiple domains as a strategy to capture more ASVs from long amplicons in diverse commmunities, and incorporating phylogenetic information into taxonomic assignments.

@@

Sampling was conducted at two sites, near the villages of Angaradebou (Ang: N 9.75456° E 2.14064°) and Gando (Gan: N 9.75678° E 2.31058°) approximately 30~km apart in the \emph{Forêt Classée de l'Ouémé Supérieur} (Upper Ouémé Forest Reserve) in central Benin.
Both sites were located in West Sudanian savannah woodlands \autocite{olson2001,yorou2014} dominated by the ECM host tree \emph{Isoberlinia doka} (Caesalpinioideae).
At each site, 25~soil samples were collected along a single 24~m linear transect at intervals of 1~m in May 2015.
One third of the sample locations (3~m spacing) were resampled one year later in June 2016, for a total of 67 samples.
For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5~cm~× 5~cm~× 5~cm was extracted with a sterilized knife blade and homogenized.
A subsample of approximately 250~mg was preserved and returned to the laboratory for extraction (see Supplementary methods).

@@

DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure~\ref{fig:regions}a) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
Due to length limitations of Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
The short amplicon (approximately 300~bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 \autocite{ihrmark2012} as the forward primer and a mix of ITS4 \autocite{white1990amplification} and ITS4a \autocite{urbina2016} as the reverse primer.
The long amplicon (approximately 1500~bp) targeted the full ITS region including the 5.8S rDNA and approximately 950~bp at the 5' end of the LSU, including the first three variable regions (Figure~\ref{fig:regions}a), using ITS1 \autocite{white1990amplification} as the forward primer and LR5 \autocite{vilgalys1990} as the reverse primer.
Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of \emph{Agaricus bisporus}.
See Supplementary methods for PCR and library preparation protocols.

@@

We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods).
We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU \autocite{michot1984}, and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure~\ref{fig:regions}a), followed by independently denoising reads from each domain, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
We used this method, implemented in the new R packages \texttt{LSUx} (\url{https://github.com/brendanf/LSUx}) and \texttt{tzara} (\url{https://github.com/brendanf/tzara}) and detailed in the Supplementary Methods, for all of the PacBio and Ion Torrent datasets.

@@

{\centering \includegraphics[width=0.7\linewidth]{transect_paper_files/figure-latex/regions-1}

@@

\caption[rDNA regions]{rDNA regions. \textbf{(a)} Partial map of rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA, and internally transcribed spacer (ITS) regions. D1--3 represent the first three variable regions in LSU, while LSU1--4 represent the conserved regions. Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces. \textbf{(b)} Total number of DADA2 ASVs vs.~fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from a set of long PacBio amplicon sequences using \texttt{LSUx}. Shorter and more conserved regions yield a greater fraction of successfully mapped reads. At a given fraction of mapped reads, more variable regions yield a greater number of unique ASVs.}\label{fig:regions}
\end{figure}

Because the \texttt{LSUx} plus \texttt{tzara} method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
The ITS2 region was extracted from the ASVs using \texttt{LSUx} for comparison to the results from the other technologies.

To account for intra-species variation and the possibility of different denoising performance between the different sequencing strategies, the pooled ITS2-ASVs from all sequencing strategies were also clustered into operational taxonomic units (OTUs) at 97\% similarity using VSEARCH v2.9.1 \autocite{rognes2016}.

\hypertarget{phylogenetic-inference-and-taxonomy-assignment}{%
\subsubsection{Phylogenetic inference and taxonomy assignment}\label{phylogenetic-inference-and-taxonomy-assignment}}

Full length long amplicon ASVs were aligned using DECIPHER \autocite{wright2015} with up to 10~iterations of alternating progressive alignment and conserved RNA secondary structure calculation, followed by 10~refinement iterations.
This alignment was truncated at a position after the D3 region corresponding to base 907 of the \emph{Saccharomyces cerevisiae} S288C reference sequence for LSU, because several sequences had introns after this position, as also observed in several fungal species by \textcite{holst-jensen1999}.
An ML tree was produced using RAxML version 8.2.12 \autocite{stamatakis2014} using the GTR+GAMMA model and rapid bootstrapping with the MRE\_IGN stopping criterion.
The tree was rooted outside kingdom Fungi and non-Fungi were removed based on taxonomic assignments (see below and Supplementary methods).

Taxonomic annotations of the RDP LSU fungal training set version 11.5 \autocite{cole2014} and Warcup ITS training set \autocite{deshpande2016} were mapped to a uniform taxonomic classification system (see Supplementary Methods).
Preliminary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP, respectively, as taxonomic references.
For each region/reference combination, taxonomy was assigned using three popular algorithms:
the RDP Naïve Bayesian Classifier \autocite[RDPC,][]{wang2007} as implemented in DADA2;
SINTAX \autocite{edgar2016a} as implemented in VSEARCH v2.9.1 \autocite{rognes2016};
and IDTAXA \autocite{murali2018}.
A relatively lax confidence threshhold of 50\% was used for all three algorithms.
Each full-length ASV was thus given up to nine preliminary taxonomic assignments (three references \(\times\) three algorithms).
ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.

For full-length long-amplicon ASVs, the preliminary taxonomic assignments were refined using the phylogenetic tree (Figure~\ref{fig:phylotax} and Supplementary Methods).
This refinement algorithm is referred to as ``PHYLOTAX'', available in the new R package \texttt{phylotax} at \url{https://github.com/brendanf/phylotax}.

ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different preliminary assignments at each rank, resulting in a consensus assignment equivalent to the ``last common ancestor'' of the preliminary assignments.
This algorithm has been used to assign a consensus taxonomy based on a list of top BLAST hits \autocite[LCAClassifier,][]{lanzen2012} or k-mer similarity scores \autocite[mothur's k-nearest neighbor method,][]{schloss2009}, but here is used to resolve conflicts between assignments from different algorithms and databases.
Strict consensus assignments were also generated for all ASVs, as a comparison to the PHYLOTAX assignments, and are referred to as ``Consensus''.

>>>>

Although ECM fungi form many well-known mushrooms (e.g., \emph{Amanita}, \emph{Cantharellus}, \emph{Boletus}), some instead produce inconspicuous (e.g., \emph{Tomentella}) or no (e.g., \emph{Cenococcum}) fruitbodies.
Even when fruitbodies are large, they are ephemeral, so study of ECM communities is facilitated by sampling of vegetative structures \autocite{horton2001}.
Unlike many saprotrophic fungi which grow easily in axenic culture, ECM fungi are usually difficult to culture, so DNA barcoding is increasingly used to investigate vegetative structures in the field.
The advent of high-throughput sequencing (HTS) has facilitated such studies by providing enough sequencing depth for metabarcoding of bulk environmental samples such as soils \autocite{Lindahl2013}.

As additional techniques and methods are developed for HTS, there is an increasing array of choices for researchers investigating fungal communities.
Fungal metabarcoding studies using short-read HTS technologies such as 454 Pyrosequencing, Illumina, and Ion Torrent
have usually targeted the rDNA internal transcribed spacer regions ITS1 or ITS2,
which are the standard molecular barcode for fungi, providing sufficient resolution to distinguish fungal species in many groups, and which are usually short enough for HTS \autocite{schoch2012,Lindahl2013}.
In some groups such as arbuscular mycorrhizal fungi, variable regions of the rDNA small subunit (SSU) are the barcode of choice \autocite{opik2010}, and variable regions of the rDNA large subunit (LSU) have also been used for barcoding \autocites[e.g.,][]{kurtzman1998,tedersoo2015,house2016}.
The resulting sequencing reads are clustered by sequence similarity to form operational taxonomic units (OTUs), which are then used as the units for further community analysis \autocite{Lindahl2013}.
If taxonomic identification is desired in order to put OTUs in a wider context and associate functional information, it has usually been performed by database searches using BLAST \autocite{altschul1990,Lindahl2013} with public databases such as GenBank \autocite{benson2013} and Unite \autocite{nilsson2019a}.
However, there is potential to improve this approach at several stages, including sequencing technology, amplicon choice, clustering, and taxonomic assignment.

Different sequencing technologies have different capabilities in terms of sequencing depth and read length, as well as differing quality profiles and potential biases \autocite{yang2013}.
The rapid development of new HTS technologies, as well as subsequent iterative improvements in sequencing chemistry and read capacity, means that the technologies used in metabarcoding studies, along with any associated biases, change frequently.
As an example, the first study using HTS metabarcoding of soil fungi was published in 2009 \autocite{buee2009} using 454 Pyrosequencing; production of 454 sequencers was subsequently discontinued in 2015, and sales of reagents stopped in 2016 \autocite{hollmer2013}.
This brings into question the comparability of studies conducted only a few years apart.
Existing studies which sequenced the same environmental samples using different HTS technologies \autocites[e.g.,][]{claesson2010,smith2014,divoll2018,Kennedy2018,speranskaya2018,Tedersoo2018} have found that most differences in results seem to be attributable to differences in sequencing depth or different primer biases, rather than differences in the technologies themselves.
Only a few of these studies have controlled for primer biases by using the same primer pairs in each technology \autocite{claesson2010,divoll2018,speranskaya2018}, and to our knowledge none have sequenced the same PCR products using multiple HTS technologies.

ITS1 and ITS2 often have suitable variation to distinguish species, although closely related species may share identical ITS sequences in certain groups such as various Pezizomycotina \autocite{schoch2012},
but this variability means that they cannot be reliably aligned over the fungal kingdom \autocite{Lindahl2013,Tedersoo2018}.
Additionally, the wide range of length variation of these regions may introduce bias in recovery of different taxa \autocite{ihrmark2012,tedersoo2015,palmer2018}.
Further bias is introduced by variation in the 5.8S region which separates the two ITS regions, as well as in the 5' end of LSU, which makes it difficult to design primers that are suitable for all fungi \autocite{tedersoo2015}.

Distance-based clustering conflates intra-species variation and sequencing error \autocite{nilsson2008,lindner2011}, and results are dataset-specific.
In contrast, more recent denoising methods such as DADA2 \autocite{callahan2017}, Deblur \autocite{amir2017}, and UNOISE2 \autocite{edgar2016} utilize read quality information to control for sequencing error while preserving intra-species variation.
The resulting units are known as amplicon sequence variants (ASVs) or exact sequence variants (ESVs), as they should represent true amplicon sequences from the sample.
Unlike cluster-based OTUs, ASVs can capture variation of as little as one base pair, although alpha and beta diversity estimates based on ASVs and OTUs at different clustering thresholds are highly correlated \autocite{glassman2018,botnen2018}.
ASVs have been suggested to be less dataset specific than cluster-based OTUs \autocite{callahan2017}.
Support for PacBio has recently been added to DADA2 \autocite{callahan2019}, but its application requires greater sequencing depth for longer reads, especially in high diversity samples.

Because both OTU clustering and denoised ASVs may ``clump'' different species into a single unit and ``split'' a single species into multiple units \autocite{ryberg2015}, diversity measures based on counting species within a community or shared species between two communities may give different results depending on the clustering threshold.
In contrast, phylogenetic community distance measures \autocite{wong2016} are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.
Phylogenetic placement algorithms have been developed to place short amplicon reads onto a reference tree \autocite{munch2008a,munch2008,matsen2010,berger2011}, but are not easy to apply to ITS sequences because they require that the query sequences be aligned to a reference alignment.
Additionally, methods exist to place OTUs on a simplified tree based on taxonomic assignments \autocite{tedersoo2018}, or to create hybrid trees using ITS and a more conserved marker such as SSU or LSU based on matching taxonomic annotations in reference databases \autocite{fouquier2016}, but these approaches are only applicable to sequences of known taxonomic affiliation.

Assignment of taxonomic identities to environmental sequences is dependent on both the reference database and the algorithm used.
Although the public INSDC databases \autocite{karsch-mizrachi2018} are often used for sequence identification, the open nature of submission to these databases results in a substantial fraction of incorrect taxonomic annotations \autocite{bidartondo2008,nilsson2006,steinegger2020} as well as sequences of poor technical quality \autocite{ashelford2006,nilsson2012}.
Consequently, taxonomic assignments based on these databases may be incorrect or inconsistent \autocite{nilsson2005}.
Several curated databases also exist which attempt to address these issues and which cover the whole fungal kingdom.
The Unite database is an attempt to include all publicly available high-quality ITS sequences (800~000 as of release 8.0), originally limited to fungi but now expanded to include all eukaryotes \autocite{nilsson2019a}, where efforts have been made to correct incorrect annotations and exclude low-quality sequences \autocite{abarenkov2018a}.
The Ribosomal Data Project \autocite[RDP,][]{cole2014} hosts two additional manually curated fungal barcode sequence databases, which are specifically intended for use in taxonomic assignment of sequences:
the Warcup ITS training set, containing 18~000 manually curated fungal ITS sequences \autocite{deshpande2016}, and the RDP fungal LSU training set, containing 8000 manually curated LSU sequences from fungi and 3000 from other eukaryotic groups \autocite[RDP-LSU,][]{liu2012}.
Although the quality of sequences and taxonomic annotations is undoubtedly higher in these more curated databases, they are inherently limited in taxonomic coverage and do not include the most recently published sequences.

@@

Recent long-read HTS technologies such as Pacific Biosciences (PacBio) Single Molecule Real Time (SMRT) sequencing enable sequencing longer amplicons which include both the ITS regions and the flanking, more highly conserved SSU and/or LSU regions.
This can improve taxonomic placement of sequences that lack close database matches and allow the alignment of metabarcoding reads for subsequent phylogenetic analysis \autocite{Tedersoo2018}.
Information from phylogenetic trees produced from long-amplicon metabarcoding has the potential to both improve taxonomic assignment and provide alternative measures of community alpha and beta diversity.
PacBio sequencing has also been shown to recover longer variants of variable-length regions such as ITS1 and ITS2, which are excluded by other technologies \autocite{castano2020}.
However, long-read technologies are currently more expensive per read compared to short-read sequencing, and so their use entails a trade-off with sequencing depth and/or sample number \autocite{Kennedy2018}.

Because of the variety of sequencing platforms and analytical pipelines which have been used in metabarcoding studies, comparisons between studies may be difficult.
Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool \autocite{legendre2012}.
Knowledge of turnover scale is important when planning studies of local diversity and its environmental correlates.
Turnover scale varies between different ecosystems and taxonomic groups,
and can be measured by the range at which a Mantel correlogram indicates significant autocorrelation,
or by fitting a function to an empirical distance-decay curve of community dissimilarity vs.~distance \autocite{legendre2012}.

We compare three different sequencing platforms (PacBio RS II, Illumina MiSeq, Ion Torrent Ion S5), long and short amplicons, three different taxonomic assignment algorithms (RDPC, SINTAX, IDTAXA)
with three different reference databases (Unite, Warcup, RDP-LSU),
and both non-phylogenetic and phylogenetic community distance measures.
We also present new algorithms for dividing the LSU into domains, combining denoising results from multiple domains as a strategy to capture more ASVs from long amplicons in diverse communities, and incorporating phylogenetic information into taxonomic assignments.
We hypothesize that (a) PacBio sequencing of the short amplicon gives less bias against longer ITS2 amplicons than Illumina and Ion Torrent, both qualitatively (recovering amplicons missed by the others) and quantitatively (greater fraction of reads in longer amplicons);
(b) our long amplicons (ITS1-LR5) recover a more complete view of the fungal community than our short amplicons (gITS7-ITS4), due to reduced length and primer biases;
(c) these differences lead to differing results for ecological metrics, specifically OTU/ASV richness and turnover distance;
and (d) incorporating LSU in the long amplicon allows for better taxonomic assignment.

@@

Sampling was conducted at two sites, near the villages of Angaradebou (Ang: N 9.75456° E 2.14064°) and Gando (Gan: N 9.75678° E 2.31058°) approximately 30~km apart in the \emph{Forêt Classée de l'Ouémé Supérieur} (Upper Ouémé Forest Reserve) in central Benin.
Both sites were located in West Sudanian savannah woodlands \autocite{olson2001,yorou2014} dominated by the ECM host tree \emph{Isoberlinia doka} (Fabaceae tr. Amherstieae).
At each site, 25~soil samples were collected along a single 24~m linear transect at intervals of 1~m in May 2015.
One third of the sample locations (3~m spacing) were resampled one year later in June 2016, for a total of 67 samples.
For each sample, coarse organic debris was removed from the soil surface and a sample of approximately 5~cm~× 5~cm~× 5~cm was extracted with an ethanol sterilized knife blade.
Each sample was sealed in a plastic zipper bag and homogenized by shaking and manually breaking apart soil aggregations.
A subsample of approximately 250~mg total of soil was collected from two locations in the homogenized soil sample and stored in a DNA preservation buffer before return to the laboratory for extraction (see Supplementary Methods \ref{Sampling-supplement} for preservation and extraction methods).

@@

DNA extracts were sequenced using four distinct strategies, with two different amplicon lengths (long and short; Figure~\ref{fig:regions}a) and three different technologies (PacBio, Ion Torrent, and Illumina) for the short amplicon.
Due to length limitations of Ion Torrent and Illumina sequencing, long amplicons were only sequenced with PacBio.
The short amplicon (approximately 300~bp) targeted the full ITS2 region as well as parts of the flanking 5.8S and large subunit (LSU) rDNA, using gITS7 \autocite{ihrmark2012} as the forward primer and an equimolar mix of ITS4 \autocite{white1990amplification} and ITS4a \autocite{urbina2016} as the reverse primer.
The long amplicon (approximately 1500~bp) targeted the full ITS region including the 5.8S rDNA and approximately 950~bp at the 5' end of the LSU, including the first three variable regions (Figure~\ref{fig:regions}a), using ITS1 \autocite{white1990amplification} as the forward primer and LR5 \autocite{vilgalys1990} as the reverse primer.
Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of \emph{Agaricus bisporus}.

For the short amplicon, forward primers included sample-specific indexes and adapters for multiplexing (Supplementary File 1).
PCR amplification was performed in 20µl reactions containing 200~µM dNTP mix, 250~µM indexed gITS7 primer,
150µM ITS4m, 2mM MgCl\textsubscript{2}, 0.1~U \emph{Taq} polymerase (Dream \emph{Taq}, Thermo Fisher Scientific, Waltham, MA, USA) and 3--7~ng purified DNA in Dream \emph{Taq} buffer.
The reaction conditions were 10~min at 95°, followed by 35~cycles of 60~s at 95°, 45~s at 56°, and 50~s at 72°, and finally 3~min at 72°.
Each reaction was conducted in three technical replicates to reduce the effect of PCR stochasticity.

For the long amplicon, both forward and reverse primers included indexes for combinatorial multiplexing (Supplementary File 2).
PCR was performed as for the short amplicons, but with 500~µM of each of the two primers.
Reaction conditions were 10~min at 95°, 30~cycles of 45~s at 95°, 45~s at 59°, and 90~s at 72°, and finally 10~min at 72°.
Each reaction was performed in three technical replicates as for short amplicons.

After pooling of technical replicates, amplicons were purified using SPRI beads \autocite{vesterinen2016} and quantified fluorometrically using Quant-iT\textsuperscript{TM} PicoGreen\textsuperscript{TM} dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.
An aliquot of 100~ng of DNA from each sample (or the total PCR product if less than 100~ng) was pooled into two libraries each for long and short amplicons.

@@

We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity, which lead to most biological variants being represented only by a cluster of reads differing by a small number of unique sequencing errors (for calculations see Supplementary Methods \ref{long-numbers}).
We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains, including the two ITS regions, 5.8S, the variable D1--3 regions of LSU \autocite{michot1984}, and the conserved LSU regions between the D regions, here referred to as LSU1--4 (Figure~\ref{fig:regions}a).
We then independently denoised reads from each domain and concatenated the denoised domains for each read.
Finally denoised reads were clustered based on 100\% ITS2 identity, and a full-length consensus ASV was calculated for each cluster.
This method, implemented in the new R packages \texttt{LSUx} (splitting reads into homologous regions; \url{https://github.com/brendanf/LSUx}) and \texttt{tzara} (reassembling regions and generating full-length consensus ASVs; \url{https://github.com/brendanf/tzara}) and detailed in Supplementary Methods \ref{denoising-supplement}, was used for all of the PacBio and Ion Torrent datasets.
Because the \texttt{LSUx} plus \texttt{tzara} method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
The ITS2 region was extracted from the ASVs using \texttt{LSUx} for comparison to the results from the other technologies.
To account for intra-species variation and the possibility of different denoising performance between the different sequencing strategies, the pooled ITS2-ASVs from all sequencing strategies were also clustered into operational taxonomic units (OTUs) at 97\% similarity using VSEARCH v2.9.1 \autocite{rognes2016}.

@@

{\centering \includegraphics[width=112mm]{transect_paper_files/figure-latex/regions-1}

@@

\caption[rDNA regions]{rDNA regions. \textbf{(a)} Partial map of rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA, and internally transcribed spacer (ITS) regions. D1--3 represent the first three variable regions in LSU, while LSU1--4 represent the conserved regions. Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces. \textbf{(b)} Total number of DADA2 ASVs vs.~fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from a set of long PacBio amplicon sequences using \texttt{LSUx}. Data from all samples were analyzed as a single pool. \emph{long}: entire long amplicon, including ITS1, 5.8S, ITS2, and partial LSU; \emph{32S} partial 32S precursor to LSU, including 5.8S, ITS2, and partial LSU; \emph{LSU}: section of LSU rDNA included in the long amplicon, from the 5' end to the LR5 primer site; \emph{ITS}: full ITS region, including ITS1, 5.8S, and ITS2. Color indicates median region length. Shorter and more conserved regions yielded a greater fraction of successfully mapped reads. At a given fraction of mapped reads, more variable regions yield a greater number of unique ASVs.}\label{fig:regions}
\end{figure}

\hypertarget{phylogenetic-inference-and-taxonomic-assignment}{%
\subsubsection{Phylogenetic inference and taxonomic assignment}\label{phylogenetic-inference-and-taxonomic-assignment}}

Full length long amplicon ASVs were aligned using DECIPHER \autocite{wright2015} with up to 10~iterations of alternating progressive alignment and conserved RNA secondary structure calculation, followed by 10~refinement iterations.
This alignment was truncated at a position after the D3 region corresponding to base 907 of the \emph{Saccharomyces cerevisiae} S288C reference sequence for LSU, because several sequences had introns after this position, as also observed in several fungal species by \textcite{holst-jensen1999}.
An ML tree was produced using RAxML version 8.2.12 \autocite{stamatakis2014} using the GTR+GAMMA model and rapid bootstrapping with the MRE\_IGN stopping criterion.
Sequences confidently (i.e., by at least five of the primary taxonomic identification methods) assigned outside kingdom Fungi were used to root the tree, and sequences outside the clade defined by confidently identified Fungi were removed (see below and Supplementary Methods \ref{rooting}).

Taxonomic annotations of the RDP-LSU training set version 11.5 \autocite{liu2012,cole2014} and Warcup ITS training set \autocite{deshpande2016} were mapped to a uniform taxonomic classification system (see Supplementary Methods \ref{reannotate}).
Primary taxonomic assignment was performed to genus level separately on the ITS region using Unite and Warcup and on the LSU region using RDP-LSU, respectively, as taxonomic references.
For each region/reference combination, taxonomy was assigned using three popular algorithms:
the RDPC \autocite{wang2007} as implemented in DADA2;
SINTAX \autocite{edgar2016a} as implemented in VSEARCH v2.9.1 \autocite{rognes2016};
and IDTAXA \autocite{murali2018}.
A relatively lax confidence threshold of 50\% was used for all three algorithms, in order to increase the amount of input for consensus algorithms.
Each full-length ASV was thus given up to nine primary taxonomic assignments (three references \(\times\) three algorithms).
ASVs from the short-amplicon datasets for which no matching long-amplicon ASV could be reconstructed were taxonomically assigned using Unite and Warcup on the full length of the short amplicon.

For full-length long-amplicon ASVs, the primary taxonomic assignments were refined based on the ML phylogenetic tree generated above using the new algorithm PHYLOTAX.
The PHYLOTAX algorithm resolves conflicts among one or more primary assignment methods using a supplied phylogenetic tree (see Figure~\ref{fig:phylotax} and Supplementary Methods \ref{phylotax}).
It is available in the new R package \texttt{phylotax} at \url{https://github.com/brendanf/phylotax}.

ASVs which were not present in the tree, either because they were not represented in the long-amplicon dataset, or because full-length ASV reconstruction failed, were given refined taxonomic assignments using a strict consensus of the different primary assignments at each rank, resulting in a consensus assignment equivalent to the ``last common ancestor'' of the primary assignments \autocite{huson2007}.
This algorithm has been used to assign a consensus taxonomy based on a list of top BLAST hits \autocites[e.g., MEGAN and LCAClassifier,][]{huson2016,lanzen2012} or k-mer similarity scores \autocite[mothur's k-nearest neighbor method,][]{schloss2009}, but here is used to resolve conflicts between assignments from different algorithms and databases.
Strict consensus assignments were also generated for all ASVs, as a comparison to the PHYLOTAX assignments, and are referred to as ``Consensus''.
